---
title: '12\.Phylogenetic Diversity - Communities'
author: "Emmi Mueller; Z620: Quantitative Biodiversity, Indiana University"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
geometry: margin=2.54cm
---

## OVERVIEW
Complementing taxonomic measures of $\alpha$- and $\beta$-diversity with evolutionary information yields insight into a broad range of biodiversity issues including conservation, biogeography, and community assembly. 
In this worksheet, you will be introduced to some commonly used methods in phylogenetic community ecology.

After completing this assignment you will know how to:

1.  incorporate an evolutionary perspective into your understanding of community ecology
2.  quantify and interpret phylogenetic $\alpha$- and $\beta$-diversity
3.  evaluate the contribution of phylogeny to spatial patterns of biodiversity

## Directions:
1. In the Markdown version of this document in your cloned repo, change "Student Name" on line 3 (above) with your name.
2. Complete as much of the worksheet as possible during class.
3. Use the handout as a guide; it contains a more complete description of data sets along with examples of proper scripting needed to carry out the exercises.
4. Answer questions in the  worksheet.
Space for your answers is provided in this document and is indicated by the ">" character.
If you need a second paragraph be sure to start the first line with ">".
You should notice that the answer is highlighted in green by RStudio (color may vary if you changed the editor theme). 
5. Before you leave the classroom today, it is *imperative* that you **push** this file to your GitHub repo, at whatever stage you are. This will enable you to pull your work onto your own computer.
6. When you have completed the worksheet, **Knit** the text and code into a single PDF file by pressing the `Knit` button in the RStudio scripting panel.
This will save the PDF output in your '8.BetaDiversity' folder.
7. After Knitting, please submit the worksheet by making a **push** to your GitHub repo and then create a **pull request** via GitHub.
Your pull request should include this file *12.PhyloCom_Worksheet.Rmd* and the PDF output of `Knitr` (*12.PhyloCom_Worksheet.pdf*).

## 1) SETUP

Typically, the first thing you will do in either an R script or an RMarkdown file is setup your environment. 
This includes things such as setting the working directory and loading any packages that you will need.

In the R code chunk below, provide the code to:  
1. clear your R environment,  
2. print your current working directory,  
3. set your working directory to your `/Week7-PhyloCom` folder,  
4. load all of the required R packages (be sure to install if needed), and  
5. load the required R source file.

```{r}
rm(list = ls())
getwd()
setwd("~/GitHub/QB2019_Mueller/2.Worksheets/12.PhyloCom")

package.list <- c('picante', 'ape', 'vegan', 'seqinr','fossil','reshape','simba')
for(package in package.list){
  if(!require(package, character.only = TRUE, quietly = TRUE)){
    install.packages(package, repos='http://cran.us.r-project.org')
    library(package, character.only = TRUE)
  }
}

source("./bin/MothurTools.R")
```

## 2) DESCRIPTION OF DATA

### need to discuss data set from spatial ecology!
 
In 2013 we sampled > 50 forested ponds in Brown County State Park, Yellowwood State Park, and Hoosier National Forest in southern Indiana. 
In addition to measuring a suite of geographic and environmental variables, we characterized the diversity of bacteria in the ponds using molecular-based approaches.
Specifically, we amplified the 16S rRNA gene (i.e., the DNA sequence) and 16S rRNA transcripts (i.e., the RNA transcript of the gene) of bacteria.
We used a program called `mothur` to quality-trim our data set and assign sequences to operational taxonomic units (OTUs), which resulted in a site-by-OTU matrix.  
In this module we will focus on taxa that were present (i.e., DNA), but there will be a few steps where we need to parse out the transcript (i.e., RNA) samples.
See the handout for a further description of this week's dataset.

## 3) LOAD THE DATA

In the R code chunk below, do the following:  
1. load the environmental data for the Brown County ponds (*20130801_PondDataMod.csv*),  
2. load the site-by-species matrix using the `read.otu()` function,  
3. subset the data to include only DNA-based identifications of bacteria,  
4. rename the sites by removing extra characters,  
5. remove unnecessary OTUs in the site-by-species, and   
6. load the taxonomic data using the `read.tax()` function from the source-code file.

```{r}
env <- read.table("data/20130801_PondDataMod.csv", sep = ",", header = TRUE)
env <- na.omit(env)

comm <- read.otu(shared = "./data/INPonds.final.rdp.shared", cutoff = "1")
comm <- comm[grep("*-DNA", rownames(comm)),]
rownames(comm) <- gsub("\\-DNA", "", rownames(comm))
rownames(comm) <- gsub("\\_", "", rownames(comm))

comm <- comm[rownames(comm) %in% env$Sample_ID, ]
comm <- comm[,colSums(comm) > 0]
tax <- read.tax(taxonomy = "./data/INPonds.final.rdp.1.cons.taxonomy")


```

Next, in the R code chunk below, do the following:  
1. load the FASTA alignment for the bacterial operational taxonomic units (OTUs),  
2. rename the OTUs by removing everything before the tab (\\t) and after the bar (|),  
3. import the *Methanosarcina* outgroup FASTA file,  
4. convert both FASTA files into the DNAbin format and combine using `rbind()`,  
5. visualize the sequence alignment,  
6. using the alignment (with outgroup), pick a DNA substitution model, and create a phylogenetic distance matrix,  
7. using the distance matrix above, make a neighbor joining tree,  
8. remove any tips (OTUs) that are not in the community data set,  
9. plot the rooted tree.

```{r}
ponds.cons <- read.alignment(file = "./data/INPonds.final.rdp.1.rep.fasta", format = "fasta")
ponds.cons$nam <- gsub("\\|.*$", "", gsub("^.*?\t", "", ponds.cons$nam))

outgroup <- read.alignment(file = "./data/methanosarcina.fasta", format = "fasta")

DNAbin <- rbind(as.DNAbin(outgroup), as.DNAbin(ponds.cons))

image.DNAbin(DNAbin, show.labels = T, cex.lab = 0.05, las = 1)

seq.dist.jc <- dist.dna(DNAbin, model = "JC", pairwise.deletion = FALSE)

phy.all <- bionj(seq.dist.jc)
phy <- drop.tip(phy.all, phy.all$tip.label[!phy.all$tip.label %in% c(colnames(comm), "Methanosarcina")])
outgroup <- match("Methanosarcina", phy$tip.label)
phy <- root(phy, outgroup, resolve.root = TRUE)
par(mar = c(1,1,2,1) + 0.1)
plot.phylo(phy, main = "Neighbor Joining Tree", "phylogram", show.tip.label = FALSE, use.edge.length = FALSE, direction = "right", cex = 0.6, label.offset = 1)
```

## 4) PHYLOGENETIC ALPHA DIVERSITY

### A.  Faith's Phylogenetic Diversity (PD)
In the R code chunk below, do the following:  
1. calculate Faith's D using the `pd()` function.

```{r}
pd <- pd(comm, phy, include.root = FALSE)
```

In the R code chunk below, do the following:  
1. plot species richness (S) versus phylogenetic diversity (PD),  
2. add the trend line, and  
3. calculate the scaling exponent. 

```{r}
par(mar = c(5,5,4,1) + 0.1)
plot(log(pd$S), log(pd$PD), pch = 20, col = "red", las = 1, xlab = "ln(S)", ylab = "ln(PD)", cex.main = 1, main = "Phylodiversity (PD) vs. Taxonomic richness (S)")

fit <- lm('log(pd$PD) ~ log(pd$S)')
abline(fit, col = "red", lw = 2)
exponent <- round(coefficients(fit)[2], 2)
legend("topleft", legend = paste("Scaling exponent =", exponent, sep = ""), bty = "n", lw = 2, col = "red")

```

***Question 1***: Answer the following questions about the PD-S pattern.    
a.  Based on how PD is calculated, why should this metric be related to taxonmic richness? 
b.  Describe the relationship between taxonomic richness and phylodiversity.
c.  When would you expect these two estimates of diversity to deviate from one another? 
d.  Interpret the significance of the scaling PD-S scaling exponent.

> ***Answer 1a***: As PD is calculated by the branch lengths between pairs of species, the more species in the sample, the more branch length to add up and the higher the PD. 
> ***Answer 1b***: Taxonomic richness is related to PD by a 3/4 scaling exponent.
> ***Answer 1c***:  If species from samples were phylogeneically clustered or overdispersed, this relationship would differ.
> ***Answer 1d***: Because phylodiversity is the length of a branching structure, it is unsurprising that the scaling exponent is 3/4ths as this is characteristic of the relationship between single dimension metric such as richness and a branching structure like phylogeny.

**i. Randomizations and Null Models**  

In the R code chunk below, do the following:  
1. estimate the standardized effect size of PD using the `richness` randomization method.  

```{r}
ses.pd.rich <- ses.pd(comm[1:2,], phy, null.model = "richness", runs = 25, include.root = FALSE)
ses.pd.rich
ses.pd.freq <- ses.pd(comm[1:2,], phy, null.model = "frequency", runs = 25, include.root = FALSE)
ses.pd.freq
ses.pd.taxa <- ses.pd(comm[1:2,], phy, null.model = "taxa.labels", runs = 25, include.root = FALSE)
ses.pd.taxa
```

***Question 2***: Using `help()` and the table above, run the `ses.pd()` function using two other null models and answer the following questions:

a.  What are the null and alternative hypotheses you are testing via randomization when calculating `ses.pd`?
b.  How did your choice of null model influence your observed ses.pd values?  Explain why this choice affected or did not affect the output. 

> ***Answer 2a***:  The null hypothesis is that our ponds are the same level of phylogenetic diversity than expected based on the null model. The alternative hypothesis is that our ponds are more phylogenetically diverse.
> ***Answer 2b***:  The only null model to produce statistically significant phylogenetic diversity from the null model was frequency as this takes into account abundances within the community that richness and taxa labels do not.

### B.  Phylogenetic Dispersion Within a Sample
Another way to assess phylogenetic $\alpha$-diversity is to look at dispersion within a sample. 

**i. Phylogenetic Resemblance Matrix**

In the R code chunk below, do the following:  
1. calculate the phylogenetic resemblance matrix for taxa in the Indiana ponds data set.

```{r}
phydist <- cophenetic.phylo(phy)
```

**ii.  Net Relatedness Index (NRI)**

In the R code chunk below, do the following:  
1. Calculate the NRI for each site in the Indiana ponds data set. 

```{r}
ses.mpd <- ses.mpd(comm, phydist, null.model = "taxa.labels", abundance.weighted = TRUE, runs = 25)

NRI <- as.matrix(-1 * ((ses.mpd[,2] - ses.mpd[,3]) / ses.mpd[,4]))
rownames(NRI) <- row.names(ses.mpd)
colnames(NRI) <- "NRI"
```

**iii.  Nearest Taxon Index (NTI)**

In the R code chunk below, do the following:
1. Calculate the NTI for each site in the Indiana ponds data set. 

```{r}
ses.mntd <- ses.mntd(comm, phydist, null.model = "taxa.labels", abundance.weighted = TRUE, runs = 25)

NTI <- as.matrix(-1 * ((ses.mntd[,2] - ses.mntd[,3]) / ses.mntd[,4]))
rownames(NTI) <- row.names(ses.mntd)
colnames(NTI) <- "NTI"
```

***Question 3***: 

a. In your own words describe what you are doing when you calculate the NRI. 
b. In your own words describe what you are doing when you calculate the NTI. 
c. Interpret the NRI and NTI values you observed for this dataset. 
d. In the NRI and NTI examples above, the arguments "abundance.weighted = FALSE" means that the indices were calculated using presence-absence data.
Modify and rerun the code so that NRI and NTI are calculated using abundance data. 
How does this affect the interpretation of NRI and NTI?

> ***Answer 3a***:  NRI compares the mean phylogenetic distance of a sample to the mean phylogenetic distance of a randomized sample to determine if the sample is more or less clustered than a randomized sample.
> ***Answer 3b***:  NTI compares the nearest taxon distance of a sample for each taxa to the nearest taxon distance of a randomized sample for each taxa.
> ***Answer 3c***:  Using NRI, all samples showed overdispersion, using NTI, most were overdispersed but some were observed to be clustered.
> ***Answer 3d***:  Using abundance weighted NRI and NTI, most samples were seen to be clustered.

## 5) PHYLOGENETIC BETA DIVERSITY 

### A. Phylogenetically Based Community Resemblance Matrix
In the R code chunk below, do the following:  
1. calculate the phylogenetically based community resemblance matrix using Mean Pair Distance, and  
2. calculate the phylogenetically based community resemblance matrix using UniFrac distance.

```{r}
dist.mp <- comdist(comm, phydist)
dist.uf <- unifrac(comm, phy)



```

In the R code chunk below, do the following:  
1. plot Mean Pair Distance versus UniFrac distance and compare.

```{r}
par(mar = c(5,5,2,1) + 0.1)
plot(dist.mp, dist.uf,
     pch = 20, col = "red", las = 1, asp = 1, xlim = c(0.15, 0.5), ylim = c(0.15, 0.5),
     xlab = "Mean Pair Distance", ylab = "UniFrac Distance")
abline(b = 1, a = 0, lty = 2)
text(0.5, 0.47, "1:1")

```

***Question 4***: 

a. In your own words describe Mean Pair Distance, UniFrac distance, and the difference between them.
b. Using the plot above, describe the relationship between Mean Pair Distance and UniFrac distance. 
Note: we are calculating unweighted phylogenetic distances (similar to incidence based measures).
That means that we are not taking into account the abundance of each taxon in each site.
c. Why might MPD show less variation than UniFrac?

> ***Answer 4a***: MPD compares the phylogenetic distance between all pairs of samples. UniFrac distance calculates the ratio of unshared to total branch lengths between two samples.
> ***Answer 4b***: There is very little variation in MPD and MPD does not explain the varition in UniFrac Distance.
> ***Answer 4c***: If PD is similar between samples but samples are phylogenetically different, MPD would show less variation than UniFrac.


### B. Visualizing Phylogenetic Beta-Diversity
Now that we have our phylogenetically based community resemblance matrix, we can visualize phylogenetic diversity among samples using the same techniques that we used in the $\beta$-diversity module from earlier in the course.

In the R code chunk below, do the following:  
1. perform a PCoA based on the UniFrac distances, and  
2. calculate the explained variation for the first three PCoA axes.

```{r}
pond.pcoa <- cmdscale(dist.uf, eig = T, k = 3)
explainvar1 <- round(pond.pcoa$eig[1] / sum(pond.pcoa$eig), 3) *100
explainvar2 <- round(pond.pcoa$eig[2] / sum(pond.pcoa$eig), 3) *100
explainvar3 <- round(pond.pcoa$eig[3] / sum(pond.pcoa$eig), 3) *100

sum.eig <- sum(explainvar1, explainvar2, explainvar3)
```

Now that we have calculated our PCoA, we can plot the results.

In the R code chunk below, do the following:  
1. plot the PCoA results using either the R base package or the `ggplot` package,  
2. include the appropriate axes,   
3. add and label the points, and    
4. customize the plot. 

```{r}
plot(pond.pcoa$points[,1], pond.pcoa$points[,2],
     xlim = c(-0.2,0.2), ylim = c(-0.16, 0.16),
     xlab = paste("PCoA 1 (", explainvar1, "%)", sep = ""),
     ylab = paste("PCoA 2(", explainvar2, "%)", sep = ""),
     pch = 16, cex = 2.0, type = "n", cex.lab = 1.5, cex.axis = 1.2, axes = FALSE)

axis(side = 1, labels = T, lwd.ticks = 2, cex.axis = 1.2, las = 1)
axis(side = 2, labels = T, lwd.ticks = 2, cex.axis = 1.2, las = 1)
abline(h= 0, v=0, lty=3)
box(lwd = 2)

points(pond.pcoa$points[,1], pond.pcoa$points[,2],
       pch = 19, cex = 3, bg = "gray", col = "gray")
text(pond.pcoa$points[,1], pond.pcoa$points[,2], labels = row.names(pond.pcoa$points))
```

In the following R code chunk:
1. perform another PCoA on taxonomic data using an appropriate measure of dissimilarity, and
2. calculate the explained variation on the first three PCoA axes. 
```{r}
pond.dist <- vegdist(comm, dist= "bray")
pond.pcoa <- cmdscale(pond.dist, eig = T, k = 3)
explainvar1 <- round(pond.pcoa$eig[1] / sum(pond.pcoa$eig), 3) *100
explainvar2 <- round(pond.pcoa$eig[2] / sum(pond.pcoa$eig), 3) *100
explainvar3 <- round(pond.pcoa$eig[3] / sum(pond.pcoa$eig), 3) *100

sum.eig <- sum(explainvar1, explainvar2, explainvar3)

plot(pond.pcoa$points[,1], pond.pcoa$points[,2],
     xlim = c(-0.2,0.2), ylim = c(-0.16, 0.16),
     xlab = paste("PCoA 1 (", explainvar1, "%)", sep = ""),
     ylab = paste("PCoA 2(", explainvar2, "%)", sep = ""),
     pch = 16, cex = 2.0, type = "n", cex.lab = 1.5, cex.axis = 1.2, axes = FALSE)

axis(side = 1, labels = T, lwd.ticks = 2, cex.axis = 1.2, las = 1)
axis(side = 2, labels = T, lwd.ticks = 2, cex.axis = 1.2, las = 1)
abline(h= 0, v=0, lty=3)
box(lwd = 2)

points(pond.pcoa$points[,1], pond.pcoa$points[,2],
       pch = 19, cex = 3, bg = "gray", col = "gray")
text(pond.pcoa$points[,1], pond.pcoa$points[,2], labels = row.names(pond.pcoa$points))
```

***Question 5***: Using a combination of visualization tools and percent variation explained, how does the phylogenetically based ordination compare or contrast with the taxonomic ordination? 
What does this tell you about the importance of phylogenetic information in this system?

> ***Answer 5***: Using Bray-Curtis distances explains much more of the variation between samples than does the Unifrac distance meaning the phylogenetic relationship between the samples is closer than the abundance distributions.
### C. Hypothesis Testing

**i. Categorical Approach**

In the R code chunk below, do the following:  
1. test the hypothesis that watershed has an effect on the phylogenetic diversity of bacterial communities.

```{r}
watershed <- env$Location
adonis(dist.uf ~ watershed, permutations = 999)
adonis(
  vegdist(
    decostand(comm, method = "log"),
    method = "bray") ~ watershed, permutations = 999
)

```

**ii. Continuous Approach**

In the R code chunk below, do the following:
1. from the environmental data matrix, subset the variables related to physical and chemical properties of the ponds, and  
2. calculate environmental distance between ponds based on the Euclidean distance between sites in the environmental data matrix (after transforming and centering using `scale()`).

```{r}
envs <- env[,5:19]
envs <- envs[, -which(names(envs) %in% c("TDS", "Salinity", "Cal_Volume"))]
env.dist <- vegdist(scale(envs), method = "euclid")

```

In the R code chunk below, do the following:  
1. conduct a Mantel test to evaluate whether or not UniFrac distance is correlated with environmental variation.

```{r}
mantel(dist.uf, env.dist)

```

Last, conduct a distance-based Redundancy Analysis (dbRDA). 

In the R code chunk below, do the following:  
1. conduct a dbRDA to test the hypothesis that environmental variation effects the phylogenetic diversity of bacterial communities,  
2. use a permutation test to determine significance, and
3. plot the dbRDA results

```{r}
ponds.dbrda <- vegan::dbrda(dist.uf ~ ., data = as.data.frame(scale(envs)))
anova(ponds.dbrda, by = "axis")
ponds.fit <- envfit(ponds.dbrda, envs, perm = 999)
ponds.fit

dbrda.explainvar1 <- round(ponds.dbrda$CCA$eig[1] / sum(c(ponds.dbrda$CCA$eig, ponds.dbrda$CCA$eig)), 3) * 100
dbrda.explainvar2 <- round(ponds.dbrda$CCA$eig[2] / sum(c(ponds.dbrda$CCA$eig, ponds.dbrda$CCA$eig)), 3) * 100

par(mar = c(5,5,4,4) + 0.1)

plot(scores(ponds.dbrda, display = "wa"), xlim = c(-2,2), ylim = c(-2,2),
     xlab = paste("dbRDA 1 (", dbrda.explainvar1, "%)", sep = ""),
     ylab = paste("dbRDA 2 (", dbrda.explainvar2, "%)", sep = ""),
     pch = 16, cex = 2, type = "n", cex.lab = 1.5, cex.axis = 1.2, axes = FALSE)

axis(side = 1, labels = T, lwd.ticks = 2, cex.axis = 1.2, las = 1)
axis(side =2, labels = T, lwd.ticks = 2, cex.axis = 1.2, las =1)
abline(h = 0, v = 0, lty = 3)
box(lwd = 2)

points(scores(ponds.dbrda, display = "wa"),
       pch = 19, cex = 3, bg = "gray", col = "gray")
text(scores(ponds.dbrda, display = "wa"),
     labels = row.names(scores(ponds.dbrda, display = "wa")), cex = 0.5)

vectors <- scores(ponds.dbrda, display = "bp")
arrows(0,0, vectors[,1] * 2, vectors[,2] *2, lwd = 2, lty = 1, length = 0.2, col = "red")
text(vectors[,1] * 2, vectors[,2]*2, pos = 3, labels = row.names(vectors))
axis(side = 3, lwd.ticks = 2, cex.axis = 1.2, las = 1, col = "red", lwd = 2.2, at = pretty(range(vectors[,1])) *2, labels = pretty(range(vectors[,1])))
axis(side = 4, lwd.ticks = 2, cex.axis = 1.2, las = 1, col = "red", lwd = 2.2, at = pretty(range(vectors[,2])) *2, labels = pretty(range(vectors[,2])))
```

***Question 6***: Based on the multivariate procedures conducted above, describe the phylogenetic patterns of $\beta$-diversity for bacterial communities in the Indiana ponds.   

> ***Answer 6***: While there is not much separation in the phylogenetic relatedness of Indiana pond samples, the differences that are seen are largely seaparted along a axis of DOC and Temperature.

## 6) SPATIAL PHYLOGENETIC COMMUNITY ECOLOGY

### A. Phylogenetic Distance-Decay (PDD)

A distance decay (DD) relationship reflects the spatial autocorrelation of community similarity.
That is, communities located near one another should be more similar to one another in taxonomic composition than distant communities.
(This is analagous to the isolation by distance (IBD) pattern that is commonly found when examining genetic similarity of a populations as a function of space.)
Historically, the two most common explanations for the taxonomic DD are that it reflects spatially autocorrelated environmental variables and the influence of dispersal limitation. 
However, if phylogenetic diversity is also spatially autocorrelated, then evolutionary history may also explain some of the taxonomic DD pattern.
Here, we will construct the phylogenetic distance-decay (PDD) relationship

First, calculate distances for geographic data, taxonomic data, and phylogenetic data among all unique pair-wise combinations of ponds.

In the R code chunk below, do the following:  
1. calculate the geographic distances among ponds,  
2. calculate the taxonomic similarity among ponds,  
3. calculate the phylogenetic similarity among ponds, and  
4. create a dataframe that includes all of the above information. 

```{r}
long.lat <- as.matrix(cbind(env$long, env$lat))
coord.dist <- earth.dist(long.lat, dist = TRUE)
bray.curtis.dist <- 1 - vegdist(comm)
unifrac.dist <- 1- dist.uf

unifrac.dist.ls <- liste(unifrac.dist, entry = "unifrac")
bray.curtis.dist.ls <- liste(bray.curtis.dist, entry = "bray.curtis")
coord.dist.ls <- liste(coord.dist, entry = "geo.dist")
env.dist.ls <- liste(env.dist, entry = "env.dist")

df <- data.frame(coord.dist.ls, bray.curtis.dist.ls[,3], unifrac.dist.ls[,3], env.dist.ls[,3])
names(df)[4:6] <- c("bray.curtis", "unifrac", "env.dist")
```
Now, let's plot the DD relationships:  
In the R code chunk below, do the following:  
1. plot the taxonomic distance decay relationship,   
2. plot the phylogenetic distance decay relationship, and  
3. add trend lines to each.

```{r}
par(mfrow = c(2,1), mar = c(1,5,2,1) + 0.1, oma = c(2,0,0,0))
plot(df$geo.dist, df$bray.curtis, xlab = "", xaxt = "n", las = 1, ylim = c(0.1, 0.9), ylab = "Bray-Curtis Similarity", main = "Distance Decay", col = "SteelBlue")
DD.reg.bc <- lm(df$bray.curtis ~ df$geo.dist)
summary(DD.reg.bc)
abline(DD.reg.bc, col = "red4", lwd = 2)

par(mar = c(2,5,1,1) + 0.1)
plot(df$geo.dist, df$unifrac, xlab = "", las = 1, ylim = c(0.1, 0.9), ylab = "Unifrac Similarity", col = "darkorchid4")
DD.reg.uni <- lm(df$unifrac ~ df$geo.dist)
summary(DD.reg.uni)
abline(DD.reg.uni, col = "red4", lwd = 2)

mtext("Geographic Distance (km)", side = 1, adj = 0.55, line = 0.5, outer = TRUE)

```

In the R code chunk below, test if the trend lines in the above distance decay relationships are different from one another.

```{r}
diffslope(df$geo.dist, df$unifrac, df$geo.dist, df$bray.curtis)

```

***Question 7***: Interpret the slopes from the taxonomic and phylogenetic DD relationships. If there are differences, hypothesize why this might be.   

> ***Answer 7***: The slopes are significantly different. In a distance decay, the members of a sample may not be the same due to dispersal limitiatons, yet community function may necessitate bacteria from a similar phylogentic distribution to be productive.


## SYNTHESIS

Ignoring technical or methodological constraints, discuss how phylogenetic information could be useful in your own research. 
Specifically, what kinds of phylogenetic data would you need? How could you use it to answer important questions in your field? 
In your response, feel free to consider not only phylogenetic approaches related to phylogenetic community ecology, but also those we discussed last week in the PhyloTraits module, or any other concepts that we have not covered in this course.

>***Synthesis***: Given that I work on microbial systems, phylogenetic information is already build into community sampling. Phylogenetic information from 16S rRNA sequencing combined with whole- genome sequencing can allow me to assess functional traits of a community in a phylogenetic context.